A Java Wrapper For Embarresingly Parallel Problems

### 2013-10-11 paper/paper.tex
Fuinished RD of paper

DISCUSS EMBARRESINGKLY PARALLEL (GET 3 LINKS BEFORE HAND)

TALK ABOUT BIOLOGICAL SCIENCES AND KAKS?

@@

BACKGROUND ON WORK WITH DAVE. INTRO TO KAKS.

Instead of attempting to rewrite the preexisting program, we decided it would be
better to look at other options. Trying to parallelize the serial KaKs 
Calculator program could present several problems: introducing bugs, breaking 
the KaKs algorithm, and the learning curve of C++ threading. We decided a 
different approach was neccessary, mostly due to the short time table to produce
results. A simple Java program was written to break up the input and spawn 
threads that would run BLANKS's KaKs Calculator\cite{blah}. When the input was 
calculated, the contents of the output files was concatenated together 
to produce a final input.

@@

\begin{thebibliography}{1}

\end{thebibliography}

>>>>

The embarresingly parallel class of problems is defined as a problem when the
computation graph is disconnected. The computation graph consists of nodes that
represent the functions or processes and the data being passed from one process
to another is displayed in the graph as edges. If the graph is disconnected, 
none of the processes need to communicate to complete the computations. In a 
practical situtation, there will be some internode communication to set up the 
problem and then again to accumulate the results back together. Also known as
pleasingly parallel problems, they can also easy be performed on a cluster of 
servers or in any kind of distributed network, due to its lack of dependencies. 
The term  embarresingly parallel was coined by Geoffrey Fox in his book, 
\underline{Parallel Programming Works!} \cite{book}

Programs in the field of Biology and Genetics often analyze long strings of
characters, whether they are DNA, RNA or protein sequences. In addition to the
large file sizes, there is a lot of processing that goes into comparing several
characters. This time adds up when there are thousands of strings to process.
However, these sequences often to not relate to each other and can be processed
independently. Using the embarresingly parallel approach, the processing time
for all of the sequences can be cut down exponentially with very minimal effort
by the programmer.

@@

Originally, I was assisting with research into different forms of the species 
\emph{Cicer arietinum}, the chikckpea. The two main forms, Kabuli and Desi, are
very different in size and color but we mostly wanted to see why the Desi is
more disease resistance and have more effective root nodule symbiosis and why
the Kabuli has a large but soft shell. These two types are both cultivated and
have been under human selection for several thousand years. We wanted to use a
Ka/Ks equation to quantify the evelution and selection of these chickpeas. We
found a Ka/Ks program writen by Zhang et al. from the Beijing Institute of 
Genomics \cite{kaks} but had three varients that we wanted to be analyzed, each 
with thousands of sequences to process. The Ka/Ks Calculator program takes
somewhere from five to thirty seconds to process one pair of sequences.

Instead of attempting to rewrite the preexisting program, we decided it would be
better to look at other options to process the sequences faster. Trying to 
parallelize the serial Ka/Ks program could present several problems: introducing
bugs, breaking the algorithm, dealing with the learning curve of C++ threading. 
We decided a  different approach was neccessary, mostly due to the short time we
had to produce results. A simple Java program was written to break up the input 
and spawn  threads that would run the Ka/Ks Calculator process. When the input 
was calculated, the contents of the output files was concatenated together to 
produce a final input.

@@

\begin{thebibliography}{1}
\bibitem{book}
Fox G., Williams R., and Messina G., ``\emph{Parallel Programming Works!},'', 
Morgan Kaufmann, May 1994
\bibitem{kaks}
Zhang Z, Li J, Zhao XQ, Wang J, Wong GK, Yu J., ``\emph{KaKs Calculator: 
calculating Ka and Ks through model selection and model averaging.},'',
Genomics Proteomics Bioinformatics, November 2006.
\bibitem{BLAST}
http://blast.ncbi.nlm.nih.gov/Blast.cgii
\end{thebibliography}
